---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zonal

<!-- badges: start -->
[![R CMD Check](https://github.com/mikejohnson51/zonal/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mikejohnson51/zonal/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

`zonal` is an in development package for intersecting vector aggregation units with large gridded data. While there are many libraries that seek to tackle this problem (see credits) we needed a library that could handle large gridded extents with multiple time layers with both many small vector units 



## Installation

You can install the development version of `zonal` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mikejohnson51/zonal")
```
## Example

This is a basic example:

```{r example}
library(zonal)
## basic example code
```

----

## Getting involved

1. Code style should attempt to follow the tidyverse style guide.
2. Please avoid adding significant new dependencies without a documented reason why.
3. Please attempt to describe what you want to do prior to contributing by submitting an issue.
4. Please follow the typical github fork - pull-request workflow.
5. Make sure you use roxygen and run Check before contributing.

----

## Open source licensing info
1. [TERMS](TERMS.md)
2. [LICENSE](LICENSE)

----

## Credits and references

Similar R packages:
 1. [exactexactr](https://github.com/isciences/exactextractr)
 2. [intersectr](https://github.com/USGS-R/intersectr)
 3. [areal](https://github.com/slu-openGIS/areal)
 4. [sf](https://github.com/r-spatial/sf)
 5. [raster](https://github.com/rspatial/raster)

